---
title: "Time series"
author: "Yucheng Wang yw6"
date: "4/14/2022"
output: pdf_document
---

# Google

## EDA

```{r}
library(fGarch)
library(forecast)
library(tseries)
google = read.csv("/Users/wyc/GOOGL_2006-01-01_to_2018-01-01.csv")
```

```{r}
head(google)
google_open = google$Open
google_open = ts(google_open,start = 0)
```

```{r}
plot(y=google_open,x=as.Date(google$Date),type = "l",main="The open price of Google")
```
The open price of google

```{r}
par(mfrow=c(2,1), mar = c(2, 4, 2, 2))
t.hat <- predict(loess(google_open ~ time(google_open), span=0.17))
plot(google_open)
lines(as.numeric(time(google_open)), t.hat, col="red", lwd=2)
plot(google_open-t.hat)
```
Loess, dipict the trend

```{r}
plot(log(google_open))
```
Take log to have a constant trend.

```{r}
plot(diff(log(google_open)))
plot(diff(log(google_open)),xlim=c(1500,1700))
```
Take one diff, seems no trend, after diff it is return of the stock. Does not seems to have any seasonal pattern. 

## Model building

```{r}
library(forecast)
acf(diff(log(google_open)))
pacf(diff(log(google_open)))
model = auto.arima(diff(log(google_open)), max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
Fit ARIMA model to the return. Hard to figure our the orders of the ARIMA model by acf, pacf, thus auto.arima. arima(1,0,3) best.

```{r}
arima_model = arima(x=diff(log(google_open)),order = c(1,0,3),method='ML')
summary(arima_model)
plot(diff(log(google_open)))
lines(diff(log(google_open))-arima_model$resid,col="red")
```
Bad result.

```{r}
qqnorm(arima_model$resid); qqline(arima_model$resid)
acf(arima_model$resid)
pacf(arima_model$resid)
hist(arima_model$resid,breaks=100,prob=TRUE)
library(MASS)
library(QRM)
fit1 <- fitdistr(arima_model$resid, "normal")
para <- fit1$estimate
curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
fit2 = fit.st(arima_model$resid)
p_val = rep(0,30)
for(i in 1:30){
  p_val[i] = Box.test(arima_model$residual^2,lag = i)[3]
}
stdret <- residuals(arima_model, standardize = TRUE)
library(PerformanceAnalytics)
chart.Histogram(stdret, methods = c("add.normal", "add.density"),
          colorset=c("gray","red","blue"))
```
qqplot, non-gaussian resid. Heteroscedasticity exist from the test. But we probabily can use arima(1,0,3) as a mean model

```{r error=FALSE, warning=FALSE, message=FALSE}
library(fGarch)
gfit1 = garchFit(~garch(1,0),data = diff(log(google_open)))
gfit2 = garchFit(~garch(1,1),data = diff(log(google_open)))
gfit3 = garchFit(~garch(2,0),data = diff(log(google_open)))
gfit4 = garchFit(~garch(2,1),data = diff(log(google_open)))
gfit5 = garchFit(~garch(1,2),data = diff(log(google_open)))
gfit6 = garchFit(~garch(2,2),data = diff(log(google_open)))

gfit7 = garchFit(~arma(1,3)+garch(1,0),data = diff(log(google_open)))
gfit8 = garchFit(~arma(1,3)+garch(1,1),data = diff(log(google_open)))
gfit9 = garchFit(~arma(1,3)+garch(2,0),data = diff(log(google_open)))
gfit10 = garchFit(~arma(1,3)+garch(2,1),data = diff(log(google_open)))
gfit11 = garchFit(~arma(1,3)+garch(1,2),data = diff(log(google_open)))
gfit12 = garchFit(~arma(1,3)+garch(2,2),data = diff(log(google_open)))


```

```{r}
summary(gfit1)
summary(gfit2)
summary(gfit3)
summary(gfit4)
summary(gfit5)
summary(gfit6)
summary(gfit7)
summary(gfit8)
summary(gfit9)
summary(gfit10)
summary(gfit11)
summary(gfit12)
```
According to the AIC(no big difference) and the significance, we find that generally we should have a ARMA model as mean, here the best model is ARMA(1,3)+GARCH(1,1), but we find that ma2, ma3, not significant. since AIC are similar, and the significance, we choose GARCH(1,1) as our best model, it is simple and reasonable.

```{r}
gfit = garchFit(~garch(1,1),data = diff(log(google_open)))
summary(gfit)
```

```{r}
par(mfrow=c(2,1))
plot(gfit, which = 1)
plot(gfit, which = 2)
```

```{r}
gfit = garch(diff(log(google_open)),order = c(1,1))
plot(predict(gfit))
```

```{r}
library(rugarch)
default_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)))
default_garch <- ugarchfit(spec = default_spec, data = diff(log(google_open)))
set.seed(135)
sim = ugarchsim(default_garch,n.sim=1000)
sim_rst = sim@simulation$seriesSim[,1]
matplot(y = sim@simulation$seriesSim[,1],x=3018:4017, type = "l",
       main = "Simulations of Open",col = "red",xlim=c(0,4020),ylab="index",xlab="difflogopen",ylim=c(-0.12,0.12))
lines(diff(log(google_open)))
```
```{r}
sim = rep(0,1000)
first = log(google_open)[3018]
for(i in 1:1000){
  if(i==1){
    sim[i]=first+sim_rst[1]
  }
  else{
  sim[i]=sim[i-1]+sim_rst[i]}
}
```

```{r}
plot(log(google_open),xlim = c(0,4020),ylim = c(4.5,8))
lines(x=3018:4017,y=sim,col="red",type="l")
```

```{r}
plot(exp(log(google_open)),xlim = c(0,4020),ylim=c(150,2700))
lines(x=3018:4017,y=exp(sim),col="red",type="l")
```

```{r}
acf((diff(log(google_open)))^2)
pacf((diff(log(google_open)))^2)
```

```{r}
model = auto.arima(diff(log(google_open))^2, max.p = 2, max.q = 2,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
garch(1,1) is better



# apple

## EDA

```{r}
apple = read.csv("/Users/wyc/AAPL_2006-01-01_to_2018-01-01.csv")
```

```{r}
head(apple)
apple_open = apple$Open
apple_open = ts(apple_open,start = 0)
```

```{r}
plot(y=apple_open,x=as.Date(apple$Date),type = "l",main="The open price of apple")
```
The open price of apple

```{r}
par(mfrow=c(2,1), mar = c(2, 4, 2, 2))
t.hat <- predict(loess(apple_open ~ time(apple_open), span=0.17))
plot(apple_open)
lines(as.numeric(time(apple_open)), t.hat, col="red", lwd=2)
plot(apple_open-t.hat)
```
Loess, dipict the trend

```{r}
plot(log(apple_open))
```
Take log to have a constant trend.

```{r}
plot(diff(log(apple_open)))
plot(diff(log(apple_open)),xlim=c(1500,1700))
```
Take one diff, seems no trend, after diff it is return of the stock. Does not seems to have any seasonal pattern. 

## Model building

```{r}
library(forecast)
acf(diff(log(apple_open)))
pacf(diff(log(apple_open)))
model = auto.arima(diff(log(apple_open)), max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
Fit ARIMA model to the return. Hard to figure our the orders of the ARIMA model by acf, pacf, thus auto.arima. arima(3,0,3) best.

```{r,warning=F}
arima_model = arima(x=diff(log(apple_open)),order = c(3,0,3),method='ML')
summary(arima_model)
plot(diff(log(apple_open)))
lines(diff(log(apple_open))-arima_model$resid,col="red")
```
Bad result.

```{r}
qqnorm(arima_model$resid); qqline(arima_model$resid)
acf(arima_model$resid)
pacf(arima_model$resid)
hist(arima_model$resid,breaks=100,prob=TRUE)
library(MASS)
library(QRM)
fit1 <- fitdistr(arima_model$resid, "normal")
para <- fit1$estimate
curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
fit2 = fit.st(arima_model$resid)
p_val = rep(0,30)
for(i in 1:30){
  p_val[i] = Box.test(arima_model$residual^2,lag = i)[3]
}
stdret <- residuals(arima_model, standardize = TRUE)
library(PerformanceAnalytics)
chart.Histogram(stdret, methods = c("add.normal", "add.density"),
          colorset=c("gray","red","blue"))
```
qqplot, non-gaussian resid. Heteroscedasticity exist from the test. But we probabily can use arima(1,0,3) as a mean model

```{r error=FALSE, warning=FALSE, message=FALSE}
library(fGarch)
gfit1 = garchFit(~garch(1,0),data = diff(log(apple_open)))
gfit2 = garchFit(~garch(1,1),data = diff(log(apple_open)))
gfit3 = garchFit(~garch(2,0),data = diff(log(apple_open)))
gfit4 = garchFit(~garch(2,1),data = diff(log(apple_open)))
gfit5 = garchFit(~garch(1,2),data = diff(log(apple_open)))
gfit6 = garchFit(~garch(2,2),data = diff(log(apple_open)))

gfit7 = garchFit(~arma(3,3)+garch(1,0),data = diff(log(apple_open)))
gfit8 = garchFit(~arma(3,3)+garch(1,1),data = diff(log(apple_open)))
gfit9 = garchFit(~arma(3,3)+garch(2,0),data = diff(log(apple_open)))
gfit10 = garchFit(~arma(3,3)+garch(2,1),data = diff(log(apple_open)))
gfit11 = garchFit(~arma(3,3)+garch(1,2),data = diff(log(apple_open)))
gfit12 = garchFit(~arma(3,3)+garch(2,2),data = diff(log(apple_open)))


```

```{r}
summary(gfit1)
summary(gfit2)
summary(gfit3)
summary(gfit4)
summary(gfit5)
summary(gfit6)
summary(gfit7)
summary(gfit8)
summary(gfit9)
summary(gfit10)
summary(gfit11)
summary(gfit12)
```
According to the AIC(no big difference) and the significance, we find that generally we should have a ARMA model as mean, here the best model is GARCH(1,1)

```{r}
gfit = garchFit(~garch(1,1),data = diff(log(apple_open)))
summary(gfit)
```

```{r}
par(mfrow=c(2,1))
plot(gfit, which = 1)
plot(gfit, which = 2)
```

```{r}
gfit = garch(diff(log(apple_open)),order = c(1,1))
plot(predict(gfit))
```
```{r}
library(rugarch)
default_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)))
default_garch <- ugarchfit(spec = default_spec, data = diff(log(apple_open)))
set.seed(116)
sim = ugarchsim(default_garch,n.sim=1000)
sim_rst = sim@simulation$seriesSim[,1]
matplot(y = sim@simulation$seriesSim[,1],x=3018:4017, type = "l",
       main = "Simulations of Open",col = "red",xlim=c(0,4020),ylab="index",xlab="difflogopen",ylim=c(-0.12,0.12))
lines(diff(log(apple_open)))
```
```{r}
sim = rep(0,1000)
first = log(apple_open)[3018]
for(i in 1:1000){
  if(i==1){
    sim[i]=first+sim_rst[1]
  }
  else{
  sim[i]=sim[i-1]+sim_rst[i]}
}
```

```{r}
plot(log(apple_open),xlim = c(0,4020),ylim = c(2,7))
lines(x=3018:4017,y=sim,col="red",type="l")
```

```{r}
plot(exp(log(apple_open)),xlim = c(0,4020),ylim=c(0,1200))
lines(x=3018:4017,y=exp(sim),col="red",type="l")
```

```{r}
acf((diff(log(apple_open)))^2)
pacf((diff(log(apple_open)))^2)
```

```{r}
model = auto.arima(diff(log(apple_open))^2, max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
garch(1,1) is better


# apple

## EDA

```{r}
apple = read.csv("/Users/wyc/AAPL_2006-01-01_to_2018-01-01.csv")
```

```{r}
head(apple)
apple_open = apple$Open
apple_open = ts(apple_open,start = 0)
```

```{r}
plot(y=apple_open,x=as.Date(apple$Date),type = "l",main="The open price of apple")
```
The open price of apple

```{r}
par(mfrow=c(2,1), mar = c(2, 4, 2, 2))
t.hat <- predict(loess(apple_open ~ time(apple_open), span=0.17))
plot(apple_open)
lines(as.numeric(time(apple_open)), t.hat, col="red", lwd=2)
plot(apple_open-t.hat)
```
Loess, dipict the trend

```{r}
plot(log(apple_open))
```
Take log to have a constant trend.

```{r}
plot(diff(log(apple_open)))
plot(diff(log(apple_open)),xlim=c(1500,1700))
```
Take one diff, seems no trend, after diff it is return of the stock. Does not seems to have any seasonal pattern. 

## Model building

```{r}
library(forecast)
acf(diff(log(apple_open)))
pacf(diff(log(apple_open)))
model = auto.arima(diff(log(apple_open)), max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
Fit ARIMA model to the return. Hard to figure our the orders of the ARIMA model by acf, pacf, thus auto.arima. arima(3,0,3) best.

```{r,warning=F}
arima_model = arima(x=diff(log(apple_open)),order = c(3,0,3),method='ML')
summary(arima_model)
plot(diff(log(apple_open)))
lines(diff(log(apple_open))-arima_model$resid,col="red")
```
Bad result.

```{r}
qqnorm(arima_model$resid); qqline(arima_model$resid)
acf(arima_model$resid)
pacf(arima_model$resid)
hist(arima_model$resid,breaks=100,prob=TRUE)
library(MASS)
library(QRM)
fit1 <- fitdistr(arima_model$resid, "normal")
para <- fit1$estimate
curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
fit2 = fit.st(arima_model$resid)
p_val = rep(0,30)
for(i in 1:30){
  p_val[i] = Box.test(arima_model$residual^2,lag = i)[3]
}
stdret <- residuals(arima_model, standardize = TRUE)
library(PerformanceAnalytics)
chart.Histogram(stdret, methods = c("add.normal", "add.density"),
          colorset=c("gray","red","blue"))
```
qqplot, non-gaussian resid. Heteroscedasticity exist from the test. But we probabily can use arima(1,0,3) as a mean model

```{r error=FALSE, warning=FALSE, message=FALSE}
library(fGarch)
gfit1 = garchFit(~garch(1,0),data = diff(log(apple_open)))
gfit2 = garchFit(~garch(1,1),data = diff(log(apple_open)))
gfit3 = garchFit(~garch(2,0),data = diff(log(apple_open)))
gfit4 = garchFit(~garch(2,1),data = diff(log(apple_open)))
gfit5 = garchFit(~garch(1,2),data = diff(log(apple_open)))
gfit6 = garchFit(~garch(2,2),data = diff(log(apple_open)))

gfit7 = garchFit(~arma(3,3)+garch(1,0),data = diff(log(apple_open)))
gfit8 = garchFit(~arma(3,3)+garch(1,1),data = diff(log(apple_open)))
gfit9 = garchFit(~arma(3,3)+garch(2,0),data = diff(log(apple_open)))
gfit10 = garchFit(~arma(3,3)+garch(2,1),data = diff(log(apple_open)))
gfit11 = garchFit(~arma(3,3)+garch(1,2),data = diff(log(apple_open)))
gfit12 = garchFit(~arma(3,3)+garch(2,2),data = diff(log(apple_open)))


```

```{r}
summary(gfit1)
summary(gfit2)
summary(gfit3)
summary(gfit4)
summary(gfit5)
summary(gfit6)
summary(gfit7)
summary(gfit8)
summary(gfit9)
summary(gfit10)
summary(gfit11)
summary(gfit12)
```
According to the AIC(no big difference) and the significance, we find that generally we should have a ARMA model as mean, here the best model is GARCH(1,1)

```{r}
gfit = garchFit(~garch(1,1),data = diff(log(apple_open)))
summary(gfit)
```

```{r}
par(mfrow=c(2,1))
plot(gfit, which = 1)
plot(gfit, which = 2)
```

```{r}
gfit = garch(diff(log(apple_open)),order = c(1,1))
plot(predict(gfit))
```
```{r}
library(rugarch)
default_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)))
default_garch <- ugarchfit(spec = default_spec, data = diff(log(apple_open)))
set.seed(116)
sim = ugarchsim(default_garch,n.sim=1000)
sim_rst = sim@simulation$seriesSim[,1]
matplot(y = sim@simulation$seriesSim[,1],x=3018:4017, type = "l",
       main = "Simulations of Open",col = "red",xlim=c(0,4020),ylab="index",xlab="difflogopen",ylim=c(-0.12,0.12))
lines(diff(log(apple_open)))
```
```{r}
sim = rep(0,1000)
first = log(apple_open)[3018]
for(i in 1:1000){
  if(i==1){
    sim[i]=first+sim_rst[1]
  }
  else{
  sim[i]=sim[i-1]+sim_rst[i]}
}
```

```{r}
plot(log(apple_open),xlim = c(0,4020),ylim = c(2,7))
lines(x=3018:4017,y=sim,col="red",type="l")
```

```{r}
plot(exp(log(apple_open)),xlim = c(0,4020),ylim=c(0,1200))
lines(x=3018:4017,y=exp(sim),col="red",type="l")
```

```{r}
acf((diff(log(apple_open)))^2)
pacf((diff(log(apple_open)))^2)
```

```{r}
model = auto.arima(diff(log(apple_open))^2, max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
garch(1,1) is better


# amazon

## EDA

```{r}
amazon = read.csv("/Users/wyc/AMZN_2006-01-01_to_2018-01-01.csv")
```

```{r}
head(amazon)
amazon_open = amazon$Open
amazon_open = ts(amazon_open,start = 0)
```

```{r}
plot(y=amazon_open,x=as.Date(amazon$Date),type = "l",main="The open price of amazon")
```
The open price of amazon

```{r}
par(mfrow=c(2,1), mar = c(2, 4, 2, 2))
t.hat <- predict(loess(amazon_open ~ time(amazon_open), span=0.17))
plot(amazon_open)
lines(as.numeric(time(amazon_open)), t.hat, col="red", lwd=2)
plot(amazon_open-t.hat)
```
Loess, dipict the trend

```{r}
plot(log(amazon_open))
```
Take log to have a constant trend.

```{r}
plot(diff(log(amazon_open)))
plot(diff(log(amazon_open)),xlim=c(1500,1700))
```
Take one diff, seems no trend, after diff it is return of the stock. Does not seems to have any seasonal pattern. 

## Model building

```{r}
library(forecast)
acf(diff(log(amazon_open)))
pacf(diff(log(amazon_open)))
model = auto.arima(diff(log(amazon_open)), max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
Fit ARIMA model to the return. Hard to figure our the orders of the ARIMA model by acf, pacf, thus auto.arima. arima(1,0,3) best.

```{r,warning=F}
arima_model = arima(x=diff(log(amazon_open)),order = c(1,0,3),method='ML')
summary(arima_model)
plot(diff(log(amazon_open)))
lines(diff(log(amazon_open))-arima_model$resid,col="red")
```
Bad result.

```{r}
qqnorm(arima_model$resid); qqline(arima_model$resid)
acf(arima_model$resid)
pacf(arima_model$resid)
hist(arima_model$resid,breaks=100,prob=TRUE)
library(MASS)
library(QRM)
fit1 <- fitdistr(arima_model$resid, "normal")
para <- fit1$estimate
curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
fit2 = fit.st(arima_model$resid)
p_val = rep(0,30)
for(i in 1:30){
  p_val[i] = Box.test(arima_model$residual^2,lag = i)[3]
}
stdret <- residuals(arima_model, standardize = TRUE)
library(PerformanceAnalytics)
chart.Histogram(stdret, methods = c("add.normal", "add.density"),
          colorset=c("gray","red","blue"))
```
qqplot, non-gaussian resid. Heteroscedasticity exist from the test. But we probabily can use arima(1,0,3) as a mean model

```{r error=FALSE, warning=FALSE, message=FALSE}
library(fGarch)
gfit1 = garchFit(~garch(1,0),data = diff(log(amazon_open)))
gfit2 = garchFit(~garch(1,1),data = diff(log(amazon_open)))
gfit3 = garchFit(~garch(2,0),data = diff(log(amazon_open)))
gfit4 = garchFit(~garch(2,1),data = diff(log(amazon_open)))
gfit5 = garchFit(~garch(1,2),data = diff(log(amazon_open)))
gfit6 = garchFit(~garch(2,2),data = diff(log(amazon_open)))

gfit7 = garchFit(~arma(1,3)+garch(1,0),data = diff(log(amazon_open)))
gfit8 = garchFit(~arma(1,3)+garch(1,1),data = diff(log(amazon_open)))
gfit9 = garchFit(~arma(1,3)+garch(2,0),data = diff(log(amazon_open)))
gfit10 = garchFit(~arma(1,3)+garch(2,1),data = diff(log(amazon_open)))
gfit11 = garchFit(~arma(1,3)+garch(1,2),data = diff(log(amazon_open)))
gfit12 = garchFit(~arma(1,3)+garch(2,2),data = diff(log(amazon_open)))


```

```{r}
summary(gfit1)
summary(gfit2)
summary(gfit3)
summary(gfit4)
summary(gfit5)
summary(gfit6)
summary(gfit7)
summary(gfit8)
summary(gfit9)
summary(gfit10)
summary(gfit11)
summary(gfit12)
```
According to the AIC(no big difference) and the significance, we find that generally we should have a ARMA model as mean, here the best model is GARCH(1,2)

```{r}
gfit = garchFit(~garch(1,2),data = diff(log(amazon_open)))
summary(gfit)
```

```{r}
par(mfrow=c(2,1))
plot(gfit, which = 1)
plot(gfit, which = 2)
```

```{r}
gfit = garch(diff(log(amazon_open)),order = c(1,1))
plot(predict(gfit))
```
```{r}
library(rugarch)
default_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,2)))
default_garch <- ugarchfit(spec = default_spec, data = diff(log(amazon_open)))
set.seed(2422224)
sim = ugarchsim(default_garch,n.sim=1000)
sim_rst = sim@simulation$seriesSim[,1]
matplot(y = sim@simulation$seriesSim[,1],x=3018:4017, type = "l",
       main = "Simulations of Open",col = "red",xlim=c(0,4020),ylab="index",xlab="difflogopen",ylim=c(-0.12,0.12))
lines(diff(log(amazon_open)))
```
```{r}
sim = rep(0,1000)
first = log(amazon_open)[3018]
for(i in 1:1000){
  if(i==1){
    sim[i]=first+sim_rst[1]
  }
  else{
  sim[i]=sim[i-1]+sim_rst[i]}
}
```

```{r}
plot(log(amazon_open),xlim = c(0,4020),ylim = c(3,10))
lines(x=3018:4017,y=sim,col="red",type="l")
```

```{r}
plot(exp(log(amazon_open)),xlim = c(0,4020),ylim=c(0,9400))
lines(x=3018:4017,y=exp(sim),col="red",type="l")
```

```{r}
acf((diff(log(amazon_open)))^2)
pacf((diff(log(amazon_open)))^2)
```

```{r}
model = auto.arima(diff(log(amazon_open))^2, max.p = 3, max.q = 3,max.order = 10, stationary = T, seasonal = F, trace = T, stepwise = F, approximation = F)
```
garch(1,1) is better




# comparing the three


```{r}
plot((google_open-mean(google_open))/sd(google_open),col="red")
lines((apple_open-mean(apple_open))/sd(apple_open),col="blue")
lines((amazon_open-mean(amazon_open))/sd(amazon_open),col="green")
```






































